---
title: "ggplot2_20190925_101_kim"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
---

# Chapter 8. ggplot2

Throughout the book, we will be creating plots using the **ggplot2** package.

```{r}
library(dplyr)
library(ggplot2)
```

**ggplot2** breaks plots into components in a way that permits beginners to create relatively complex and aesthetically pleasing plots using syntax that is intuitive and comparatively easy to remember.

## 8.1. The components of a graph

Data visualization shows us pretty much all the information in the data table. 
The first step in learning **ggplot2** is to be able to break a graph apart into components. 

```{r}
library(dslabs)
data(murders)
```

## 8.2. ```ggplot``` objects

The first step in creating a **ggplot2** graph is to define a ```ggplot``` object. We do this with the function ```ggplot```, which initializes the graph. The first argument is used to specity what data is associated with this object.

```{r}
ggplot(data = murders)
```

We can also pipe the data in as the first argument:

```{r}
murders %>% ggplot()
```

It rendersa plot, in this case a blank slate since no geometry has been defined. The only style choice we see is a grey background.

We can assign our plot to an object:

```{r}
p <- ggplot(data = murders)
class(p)
print(p)
```

## 8.3. Geometries

In ```ggplot2``` we create graphs by adding *layers*. Layers can define geometries, compute summary statistics, define what scales to use, or even change styles. To add layers, we use the symbol ```+```. 

> DATA %>% ```ggplot()``` + LAYER 1 + LAYER 2 + ... + LAYER N

Usually, the first added layer defines the geometry. Our possible geometries are scatterplot, barplot, histogram, smooth densities, qqplot, and boxplot.

Geometry function names follow the pattern : ```geom_X``` where X is the name of the geometry. Some examples include ```geom_point```, ```geom_bar``` and ```geom_histogram```.

For ```geom_point``` to run properly we need to provide data and a mapping. 

## 8.4. Aesthetic mappings

Aesthetic mappings describe how properties of the data connect with features of the graph, such as distance along an axis, size or color. The ```aes``` function connects data with what we see on the graph by defining aesthetic mappings, and will be one of the functions you use most often when plotting.

The outcome of the ```aes``` function is often used as the argument of a geometry function. 

```{r}
murders %>% ggplot() +
  geom_point(aes(x = population/10^6, y = total))
```

We can drop the ```x=``` and ```y=``` if we wanted. Onstead of defining plot from scratch, we can also add a layer to the ```p``` object that was defined above as ```p <- ggplot(data = murders)```:

```{r}
p + geom_point(aes(population/10^6, total))
```

Like ```dplyr``` functions, ```aes``` also uses the variable names from the object component : we can use ```population``` and ```total``` without having to call them using accessor ```$```. 

## 8.5. Layers

A second layer in the plot involves adding a label to each point to identify the state.

The ```geom_label``` and ```geom_text``` functions permit us to add text to the plot with and without a rectangle behind the text respectively.

Because each point has a label, we need an aesthetic mapping to make the connection between points and labels. 

```{r}
p + geom_point(aes(population/10^6, total)) +
  geom_text(aes(population/10^6, total, label = abb))
```

> 8.5.1. Tinkering with arguments

Each geometry function has many arguments other than ```aes``` and ```data```. They tend to be specific to the function. 

```{r}
p + geom_point(aes(population/10^6, total), size = 3) +
  geom_text(aes(population/10^6, total, label = abb))
```

```size``` is not a mapping : whereas mappings use data from specific observations and need to be inside ```aes()```, operations we want to affect all the points the same way do not need to be included inside ```aes``. 

```nudge_x``` argument moves the text slightly to the right or to the left:

```{r}
p + geom_point(aes(population/10^6, total), size = 3) +
  geom_text(aes(population/10^6, total, label = abb), nudge_x = 1)
```

## 8.6. Global versus local aesthetic mappings

If we define a mapping in ```ggplot```, all the geometries that are added as layers will default to the mapping. We redefine p :
```{r}
p <- murders %>% ggplot(aes(population / 10^6, total, label = abb))
```

and we can simply write :
```{r}
p + geom_point(size = 3) +
  geom_text(nudge_x = 1.5)
```

We can override the global mapping by defining a new mapping within each layer. These local definitions override the global:

```{r}
p + geom_point(size = 3)+
  geom_text(aes(x = 10, y = 800, label = "Hello there!"))
```

## 8.7. Scales

Our desired scales are in log-scale. 

```{r}
p + geom_point(size = 3) +
  geom_text(nudge_x = 0.05) +
  #scale_x_log10()
  scale_x_continuous(trans = "log10") +
  #scale_y_log10()
  scale_y_continuous(trans = "log10")
```

## 8.8. Labels and titles

```{r}
p + geom_point(size = 3) +  
  geom_text(nudge_x = 0.05) + 
  scale_x_log10() +
  scale_y_log10() +
  xlab("Populations in millions (log scale)") + 
  ylab("Total number of murders (log scale)") +
  ggtitle("US Gun Murders in 2010")
```

## 8.9. Categories as colors

We can change the color of the points using the ```col``` argument in the ```geom_point``` function.

```{r}
p <-  murders %>% ggplot(aes(population/10^6, total, label = abb)) +   
  geom_text(nudge_x = 0.05) + 
  scale_x_log10() +
  scale_y_log10() +
  xlab("Populations in millions (log scale)") + 
  ylab("Total number of murders (log scale)") +
  ggtitle("US Gun Murders in 2010") +
  geom_point(size = 3, color ="blue")

p
```

If we assign a categorical variable to color, it automatically assigns a different color to each category and also adds a legend.

```{r}
p + geom_point(aes(col=region), size = 3)
```

## 8.10. Annotation, shapes, and adjustments

We can add shapes or annotation to figures that are not derived directly from the aesthetic mapping ; examples include labels, boxes, shaded areas and lines. To add a line we use the ```geom_abline``` function. ggplot2 uses ```ab``` in the name to remind us we are supplying the intercept and slope.

```{r}
r <- murders %>%
  summarize(rate = sum(total) / sum(population) * 10^6) %>%
  pull(rate)
p + geom_point(aes(col = region), size = 3) +
  geom_abline(intercept = log10(r), lty = 2, color = "darkgrey") +
  geom_point(aes(col=region), size = 3)
```

```geom_abline``` doesn't use any information from the data object,

## 8.11. Add-on packages

The style of a ggplot2 graph can be changed using the ```theme``` functions. In this book we use a function in the dslabs package that automatically sets a default theme:

```{r}
ds_theme_set()
```

Many other themes are added by the package ggthemes. 

```{r}
#install.packages("ggthemes")
library(ggthemes)
p + theme_economist()
```

## 8.12. Putting it all together

```{r}
#install.packages("ggrepel")

library(ggthemes)
library(ggrepel)

r <- murders %>% 
  summarize(rate = sum(total) /  sum(population) * 10^6) %>%
  pull(rate)

murders %>% ggplot(aes(population/10^6, total, label = abb)) +   
  geom_abline(intercept = log10(r), lty = 2, color = "darkgrey") +
  geom_point(aes(col=region), size = 3) +
  geom_text_repel() + 
  scale_x_log10() +
  scale_y_log10() +
  xlab("Populations in millions (log scale)") + 
  ylab("Total number of murders (log scale)") +
  ggtitle("US Gun Murders in 2010") + 
  scale_color_discrete(name = "Region") +
  theme_economist()
```

## 8.13. Quick plots with ```qplot```

The ```qplot``` function sacrifices the flexibility provided by the ```ggplot``` approach, but allows us to generate a plot quickly.

```{r}
data(murders)
x <- log10(murders$population)
y <- murders$total

qplot(x, y)
```

## 8.14. Grids of plots

The gridExtra package permits us to graph plots next to each other.

```{r}
#install.packages("gridExtra")

library(gridExtra)

p1 <- murders %>%
  mutate(rate = total/population*10^5) %>%
  filter(population < 2*10^6) %>%
  ggplot(aes(population/10^6, rate, label = abb)) +
  geom_text() +
  ggtitle("Small States")

p2 <- murders %>%
  mutate(rate = total/population*10^5) %>%
  filter(population > 10*10^6) %>%
  ggplot(aes(population/10^6, rate, label = abb)) +
  geom_text() +
  ggtitle("Large States")

grid.arrange(p1, p2, ncol = 2)
```

## 8.15. Exercises

```{r}
library(dplyr)
library(ggplot2)
library(dslabs)
data(heights)
data(murders)
```

1. With ggplot2 plots can be saved as objects. For example we can associate a dataset with a plot object like this

```{r}
p <- ggplot(data = murders)
```

Because ```data``` is the first argument we don't need to spell it out

```{r}
p <- ggplot(murders)
```

and we can also use the pipe:

```{r}
p <- murders %>% ggplot()
```

What is class of the project ```p```?

My answer : ggplot

```{r}
class(p)
```

2. Remember that to print an object you can use the command print or simply type the object. Print the object p defined in exercise one and describe what you see.

```{r}
print(p)
```

My answer : B

3. Using the pipe %>%, create an object p but this time associated with the heights dataset instead of the murders dataset.

My answer
```{r}
p <- heights %>% ggplot()
```

4. What is the class of the object ```p``` you have just created?

My answer : ggplot

5. Now we are going to add a layers and the corresponding aesthetic mappings. For the murders data we plotted total murders versus population sizes. Explore the ```murders``` data frame to remind yourself what are the names for these two variables and select the correct answer. 

My answer : C

6. To create the scatter plot we add a layer with ```geom_point```. The aesthetic mappings require us to define the x-axis and y-axis variables respectively. So the code looks like this:

My answer
```{r}
murders %>% ggplot(aes(x = population/10^6, y = total)) +
  geom_point()
head(murders)
```

7. Note that if we don’t use argument names, we can obtain the same plot by making sure we enter the variable names in the right order like this:

```{r}
murders %>% ggplot(aes(population, total)) +
  geom_point()
```

Remake the plot but now with total in the x-axis and population in the y-axis.

My answer
```{r}
murders %>% ggplot(aes(total, population)) +
  geom_point()
```

8. If instead of points we want to add text, we can use the ```geom_text()``` or ```geom_label()``` geometries. The following code

```{r}
murders %>% ggplot(aes(population, total)) +
  geom_label()
```

will give us the error message. Why is this?

My answer : A

9. Rewrite the code above to abbreviation as the label through aes

My answer

```{r}
murders %>% ggplot(aes(population, total)) +
  geom_label(aes(label = abb))
```

10. Change the color of the labels through blue. How will we do this?

My answer : D

11. Rewrite the code above to make the labels blue.

My answer

```{r}
murders %>% ggplot(aes(population, total)) +
  geom_label(aes(label = abb), color = "blue")
```

12. Now suppose we want to use color to represent the different regions. In this case which of the following is most appropriate:

My answer : B

13. Rewrite the code above to make the labels’ color be determined by the state’s region.

My answer

```{r}
murders %>% ggplot(aes(population, total)) +
  geom_label(aes(label = abb, color = abb))
```

14. Now we are going to change the x-axis to a log scale to account for the fact the distribution of population is skewed. Let’s start by define an object p holding the plot we have made up to now

```{r}
p <- murders %>% 
  ggplot(aes(population, total, label = abb, color = region)) +
  geom_label() 
```

To change the y-axis to a log scale we learned about the scale_x_log10() function. Add this layer to the object p to change the scale and render the plot

My answer

```{r}
p <- murders %>% 
  ggplot(aes(population, total, label = abb, color = region)) +
  geom_label() +
  scale_x_log10()

p
```

15. Repeat the previous exercise but now change both axes to be in the log scale.

My answer

```{r}
p <- murders %>% 
  ggplot(aes(population, total, label = abb, color = region)) +
  geom_label() +
  scale_x_log10() +
  scale_y_log10()

p
```

16. Now edit the code above to add the title “Gun murder data” to the plot. Hint: use the ggtitle function.

My answer

```{r}
p <- murders %>% 
  ggplot(aes(population, total, label = abb, color = region)) +
  geom_label() +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Gun murder data")

p
```














