---
title: "principles_20191007_101_kim"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document:
    latex_engine: xelatex 
---

## Chapter 11. Data visualizaton principles

We aim to provide some general principles for effective data visualization. We show some examples of plot styles we should avoid, explain how to improve them, and use these as motivation for a list of principles. We compare and contrast plots that follow these principles to those that don't.

The principles are mostly based on research related to how humans detect patterns and make visual comparisons. The preferred approaches are those that best fit the way our brains process visual information. When deciding on a visualization approach, it is also important to keep our goal in mind.

We may be comparing a viewable number of quantities, describing distributions for categories or numeric values, comparing the data from two groups, or describing the relationship between two variables.

As a final note, we want to emphasize that for a data scientist it is important to **adapt and optimize graphs to the audience**. 

```{r}
library(tidyverse)
library(dslabs)
library(gridExtra)
```

### 11.1. Encoding data using visual cues

**Pie charts are a very bad way of displaying information.** The eye is good at judging linear measures and bad at judging relative areas. A bar chart or dot chart is a preferable way of displaying this type of data. 

The preffered way instead of using pie chart is to use length and position as visual cues, since humans are much better at judging linear measures. 

The barplot uses this approach by using bars of length proportional to the quantities of interest. By adding horizontal lines at strategically chosen values, in this case at every multiple of 10, we ease the visual burden of quantifying through the position of the top of the bars. 

If for some reason you need to make a pie chart, label each pie slice with its respective percentage so viewers do not have to infer them from the angles or area.

### 11.2. Know when to include 0

When using barplots, it is misinformative not to start the bars at 0. By avoiding 0, relatively small differences can be made to look much bigger than they actually are. This approach is often used when trying to exaggerate a difference. 

When using position rather than length, it is then not necessary to include 0. When plot includes 0, the space between 0 and real data adds no information and makes it harder to compare the between and within group variability.

### 11.3. Do not distort quantities

**ggplot2** defaults to using area rather than radius. Of course, we really shouldn't be using area at all since we can use position and length.

### 11.4. Order categories by a meaningful value

When one of the axes is used to show categories, as is done in barplots, the default ggplot2 behavior is to order the categories alphabetically **when they are defined by character strings**. If they are defined by factors, they are ordered by the factor levels. We rarely want to use alphabetical order. Instead, we should order by a meaningful quantity. In all the cases above, the barplots were ordered by the values being displayed. The exception was the graph showing barplots comparing browsers. In this case, we kept the order the same across the barplots to ease the comparison.

We previously learned how to use the ```reorder``` function, which helps us achieve this goal. To appreciate how the right order can help convey a message, suppose we want to create a plot to compare the murder rate across states. We are particularly interested in the most dangerous and safest states. Note the difference when we order alphabetically (the default) versus when we order by the actual rate:

```{r}
data(murders)
murders %>% mutate(murder_rate = total / population * 100000) %>%
  mutate(state = reorder(state, murder_rate)) %>%
  ggplot(aes(state, murder_rate)) +
  geom_bar(stat="identity") +
  coord_flip() +
  theme(axis.text.y = element_text(size = 6)) +
  xlab("")
```

### 11.5. Show the data

We now shift our attention to displaying data, with a focus on comparing groups.

show the data : **ggplot2** code already generates a more informative plot than the barplot by simply showing all the data points:

```{r}
heights %>% ggplot(aes(sex, height)) +
  geom_point()
```

This plot gives us an idea of the range of the data. However, this plot has limitations as well, since we can’t really see all the 238 and 812 points plotted for females and males respectively, and many points are plotted on top of each other. As we have previously described, visualizing the distribution is much more informative. But before doing this, we point out two ways we can improve a plot showing all the points.

The first is to add *jitter*, which adds a small random shift to each point. In this case, adding horizontal jitter does not alter the interpretation, since the point heights do not change, but we minimize the number of points that fall on top of each other and, therefore, get a better visual sense of how the data is distributed.

A second improvement comes from using alpha blending: making the points somewhat transparent. The more points fall on top of each other, the darker the plot, which also helps us get a sense of how the points are distributed.

```{r}
heights %>% 
  ggplot(aes(sex, height)) +
  geom_jitter(width = 0.1, alpha = 0.2) 
```

### 11.6. Ease comparisons

> 11.6.1. Use common axes

If there's so many points, it's more effective to use histograms to show distributions rather than individual points.

An important principle is to **keep the axes the same** when comparing data across to plots.

> 11.6.2. Align plots vertically to see horizontal changes and horizontaly to see vertical changes

```{r}
heights %>% 
  ggplot(aes(height, ..density..)) +
  geom_histogram(binwidth = 1, color="black") +
  facet_grid(sex~.)
```

In these histograms, the visual cue related to decreases or increases in height are shifts to the left or right respectively: horizontal changes. Aligning the plots vertically helps us see this change when the axes are fixed.

If , we want the more compact summary provided by boxplots, we then align them horizontally since, by default, boxplots move up and down with changes in height.

```{r}
 heights %>% 
  ggplot(aes(sex, height)) + 
  geom_boxplot(coef=3) + 
  geom_jitter(width = 0.1, alpha = 0.2) +
  ylab("Height in inches")
```

> 11.6.3. Consider transformations

We have motivated the use of the log transformation in cases where the changes are multiplicative. Population size was an example in which we found a log transformation to yield a more informative transformation.

The combination of an incorrectly chosen barplot and a failure to use a log transformation when one is merited can be particularly distorting.

The transformations you should consider are the logistic transformation (```logit```), useful to better see fold changes in odds, and the square root transformation (```sqrt```), useful for count data.

> 11.6.4. Visual cues to be compared should be adjacent

It is much easier to make the comparison between 1970 and 2010 for each continent when the boxplots for that continent are next to each other.

> 11.6.5. Use color

The comparison becomes even easier to make if we use color to denote the two things we want to compare.

### 11.7. Think of the color blind

About 10% of the population is color blind. Unfortunately, the default colors used in ggplot2 are not optimal for this group. However, ggplot2 does make it easy to change the color palette used in the plots.

```{r}
color_blind_friendly_cols <- 
  c("#999999", "#E69F00", "#56B4E9", "#009E73", 
    "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### 11.8. Plots for two variables

In general, you should use scatterplots to visualize the relationship between two variables. However, there are some exceptions and we describe two alternative plots here: the slope chart and the Bland-Altman plot.

> 11.8.1. Slope charts

One exception where another type of plot may be more informative is when you are comparing variables of the same type, but at different time points and for a relatively small number of comparisons. For example, comparing life expectancy between 2010 and 2015. In this case, we might recommend a slope chart.

There is no geometry for slope charts in ggplot2, but we can construct one using ```geom_line```. We need to do some tinkering to add labels. Below is an example comparing 2010 to 2015 for large western countries:

```{r}
west <- c("Western Europe","Northern Europe","Southern Europe",
          "Northern America","Australia and New Zealand")

dat <- gapminder %>% 
  filter(year%in% c(2010, 2015) & region %in% west & 
           !is.na(life_expectancy) & population > 10^7) 

dat %>%
  mutate(location = ifelse(year == 2010, 1, 2), 
         location = ifelse(year == 2015 & 
                             country %in% c("United Kingdom","Portugal"),
                           location+0.22, location),
         hjust = ifelse(year == 2010, 1, 0)) %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(year, life_expectancy, group = country)) +
  geom_line(aes(color = country), show.legend = FALSE) +
  geom_text(aes(x = location, label = country, hjust = hjust), 
            show.legend = FALSE) +
  xlab("") + 
  ylab("Life Expectancy")
```

An advantage of the slope chart is that it permits us to quickly get an idea of changes based on the slope of the lines. Although we are using angle as the visual cue, we also have position to determine the exact values. Comparing the improvements is a bit harder with a scatterplot.

In the scatterplot, we have followed the principle use common axes since we are comparing these before and after. However, if we have many points, slope charts stop being useful as it becomes hard to see all the lines.

> 11.8.2. Bland-Altman plot

Since we are primarily interested in the difference, it makes sense to dedicate one of our axes to it. The Bland-Altman plot, also know as the Tukey mean-difference plot and the MA-plot, shows the difference versus the average:

```{r}
library(ggrepel)
dat %>% 
   mutate(year = paste0("life_expectancy_", year)) %>%
   select(country, year, life_expectancy) %>% spread(year, life_expectancy) %>% 
  mutate(average = (life_expectancy_2015 + life_expectancy_2010)/2,
         difference = life_expectancy_2015 - life_expectancy_2010) %>%
  ggplot(aes(average, difference, label = country)) + 
  geom_point() +
  geom_text_repel() +
  geom_abline(lty = 2) +
  xlab("Average of 2010 and 2015") + 
  ylab("Difference between 2015 and 2010")
```

Here, by simply looking at the y-axis, we quickly see which countries have shown the most improvement. We also get an idea of the overall value from the x-axis.

### 11.9. Encoding a third variable

We encode categorical variables with color and shape. These shapes can be controlled with ```shape``` argument. 

For continuous variables, we can use color, intensity or size. We now show an example of how we do this with a case study.

When selecting colors to quantify a numeric variable, we choose between two options: sequential and diverging. Sequential colors are suited for data that goes from high to low. High values are clearly distinguished from low values. Here are some examples offered by the package ```RColorBrewer```:

```{r}
library(RColorBrewer)
display.brewer.all(type="seq")
```

Diverging colors are used to represent values that diverge from a center. We put equal emphasis on both ends of the data range: higher than the center and lower than the center. An example of when we would use a divergent pattern would be if we were to show height in standard deviations away from the average. Here are some examples of divergent patterns:

```{r}
library(RColorBrewer)
display.brewer.all(type="div")
```

### 11.10. Avoid pseudo-three-dimensional plots

Humans are not good at seeing in three dimensions and our limitation is even worse with regard to pseudo-three-dimensions. Pseudo-3D is sometimes used completely gratuitously: plots are made to look 3D even when the 3rd dimension does not represent a quantity. This only adds confusion and makes it harder to relay your message.

### 11.11. Avoid too many significant digits

By default, statistical software like R returns many significant digits. The default behavior in R is to show 7 significant digits. That many digits often adds no information and the added the visual clutter can make it hard for the viewer to understand the message.

In this case, two significant figures is more than enough and clearly makes the point that rates are decreasing:

Useful ways to change the number of significant digits or to round numbers are ```signif``` and ```round```. You can define the number of significant digits globally by setting options like this: ```options(digits = 3)```.

### 11.12. Know your audience

Graphs can be used for 1) our own exploratory data analysis, 2) to convey a message to experts, or 3) to help tell a story to a general audience. Make sure that the intended audience understands each element of the plot.

As a simple example, consider that for your own exploration it may be more useful to log-transform data and then plot it. However, for a general audience that is unfamiliar with converting logged values back to the original measurements, using a log-scale for the axis instead of log-transformed values will be much easier to digest.

### 11.13. Exercises

1. Pie charts are appropriate: D. Never. Barplots and tables are always better.

2. What is the problem with the plot below: B. The axis does not start at 0. Judging by the length, it appears Trump received 3 times as many votes when, in fact, it was about 30% more.

3. Take a look at the following two plots. They show the same information: 1928 rates of measles across the 50 states. Which plot is easier to read if you are interested in determining which are the best and worst states in terms of rates, and why? : C. The plot on the right is better because alphabetical order has nothing to do with the disease and by ordering according to actual rate, we quickly see the states with most and least rates.

4. To make the plot on the left, we have to reorder the levels of the states’ variables.

```{r}
dat <- us_contagious_diseases %>%  
  filter(year == 1967 & disease=="Measles" & !is.na(population)) %>%
  mutate(rate = count / population * 10000 * 52 / weeks_reporting)
```

Note what happens when we make a barplot:

```{r}
dat %>% ggplot(aes(state, rate)) +
  geom_bar(stat="identity") +
  coord_flip() 
```

Define these objects:
```{r}
state <- dat$state
rate <- dat$count/dat$population*10000*52/dat$weeks_reporting
```

Redefine the state object so that the levels are re-ordered. Print the new object state and its levels so you can see that the vector is not re-ordered by the levels.

My answer:
```{r}
dat %>% ggplot(aes(state, rate)) +
  geom_bar(stat="identity") +
  coord_flip() 
```

5. Now with one line of code, define the dat table as done above, but change the use mutate to create a rate variable and reorder the state variable so that the levels are reordered by this variable. Then make a barplot using the code above, but for this new dat.

My answer:
```{r}
dat <- us_contagious_diseases %>%  
  filter(year == 1967 & disease=="Measles" & !is.na(population)) %>%
  mutate(state = reorder(state, rate))

dat %>% ggplot(aes(state, rate)) +
  geom_bar(stat="identity") +
  coord_flip()
```

6. Say we are interested in comparing gun homicide rates across regions of the US. We see this plot:

```{r}
library(dslabs)
data("murders")
murders %>% mutate(rate = total/population*100000) %>%
  group_by(region) %>%
  summarize(avg = mean(rate)) %>%
  mutate(region = factor(region)) %>%
  ggplot(aes(region, avg)) +
  geom_bar(stat="identity") +
  ylab("Murder Rate Average")
```

and decide to move to a state in the western region. What is the main problem with this interpretation? : C. It does not show all the data. We do not see the variability within a region and it’s possible that the safest states are not in the West.

7. Make a boxplot of the murder rates defined as
```{r}
data("murders")
murders %>% mutate(rate = total/population*100000)
```
by region, showing all the points and ordering the regions by their median rate.

My answer:
```{r}
murders %>% ggplot(aes(region, rate)) +
  geom_boxplot(coef = 3) +
  geom_jitter(width = 0.1, alpha = 0.2)
```

8. The plots below show three continous variables. The line x=2 appears to separate the points. But it is actually not the case, which we can see by plotting the data in a couple of two dimensional points. Why is this happening? : A. Humans are not good at reading pseudo-3D plots.

### 11.14. Case study : impact of vaccines on battling infectious diseases

```{r}
library(tidyverse)
library(dslabs)
data(us_contagious_diseases)
str(us_contagious_diseases)
```

We create a temporary object ```dat``` that stores only the measles data, includes a per 100,000 rate, orders states by average value of disease and removes Alaska and Hawaii since they only became states in the late 50s. Note that there is a ```weeks_reporting``` column that tells us for how many weeks of the year data was reported. We have to adjust for that value when computing the rate.

```{r}
the_disease <- "Measles"
dat <- us_contagious_diseases %>%
  filter(!state%in%c("Hawaii","Alaska") & disease == the_disease) %>%
  mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>% 
  mutate(state = reorder(state, rate)) 
```

We can now easily plot disease rates per year. Here are the measles data from California:

```{r}
dat %>% filter(state == "California" & !is.na(rate)) %>%
  ggplot(aes(year, rate)) +
  geom_line() + 
  ylab("Cases per 10,000")  + 
  geom_vline(xintercept=1963, col = "blue")
```

We add a vertical line at 1963 since this is when the vaccine was introduced. Now can we show data for all states in one plot? We have three variables to show: year, state and rate. In the WSJ figure, they use the x-axis for year, the y-axis for state and color hue to represent rates. However, the color scale they use, which goes from yellow to blue to green to orange to red, can be improved.

In our example, we want to use a sequential palette since there is no meaningful center, just low and high rates.

We use the geometry ```geom_tile``` to tile the region with colors representing disease rates. We use a square root transformation to avoid having the really high counts dominate the plot. Notice that missing values are shown in grey. Note that once a disease was pretty much eradicated, some states stopped reporting cases all together. This is why we see so much grey after 1980.

```{r}
dat %>% ggplot(aes(year, state, fill = rate)) +
  geom_tile(color = "grey50") +
  scale_x_continuous(expand=c(0,0)) +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "Reds"), trans = "sqrt") +
  geom_vline(xintercept=1963, col = "blue") +
  theme_minimal() +  theme(panel.grid = element_blank()) +
  ggtitle(the_disease) + 
  ylab("") + 
  xlab("")

```

This plot makes a very striking argument for the contribution of vaccines. However, one limitation of this plot is that it uses color to represent quantity, which we earlier explained makes it harder to know exactly how high values are going. Position and lengths are better cues. If we are willing to lose state information, we can make a version of the plot that shows the values with position. We can also show the average for the US, which we compute like this:

```{r}
avg <- us_contagious_diseases %>%
  filter(disease==the_disease) %>% group_by(year) %>%
  summarize(us_rate = sum(count, na.rm=TRUE)/sum(population, na.rm=TRUE)*10000)
```

Now to make the plot we simply use the ```geom_line``` geometry:

```{r}
dat %>% 
  filter(!is.na(rate)) %>%
    ggplot() +
  geom_line(aes(year, rate, group = state),  color = "grey50", 
            show.legend = FALSE, alpha = 0.2, size = 1) +
  geom_line(mapping = aes(year, us_rate),  data = avg, size = 1, color = "black") +
  scale_y_continuous(trans = "sqrt", breaks = c(5, 25, 125, 300)) + 
  ggtitle("Cases per 10,000 by state") + 
  xlab("") + 
  ylab("") +
  geom_text(data = data.frame(x = 1955, y = 50), 
            mapping = aes(x, y, label="US average"), 
            color="black") + 
  geom_vline(xintercept=1963, col = "blue") 
```

In theory, we could use color to represent the categorical value state, but it is hard to pick 50 distinct colors.

### 11.15. Exercises

1. Reproduce the image plot we previously made but for smallpox. For this plot, do not include years in which cases were not reported in 10 or more weeks.

```{r}
dat <- us_contagious_diseases %>%
  filter(!state%in%c("Hawaii","Alaska") & disease == "Smallpox" & weeks_reporting > 42) %>%
  mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>% 
  mutate(state = reorder(state, rate)) 

dat %>% ggplot(aes(year, state, fill = rate)) +
  geom_tile(color = "grey50") +
  scale_x_continuous(expand=c(0,0)) +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "Reds"), trans = "sqrt") +
  theme_minimal() +  theme(panel.grid = element_blank()) +
  ggtitle("Smallpox") + 
  ylab("") + 
  xlab("")
```

2. Now reproduce the time series plot we previously made, but this time following the instructions of the previous question.

```{r}
avg <- us_contagious_diseases %>%
  filter(disease=="Smallpox") %>% group_by(year) %>%
  summarize(us_rate = sum(count, na.rm=TRUE)/sum(population, na.rm=TRUE)*10000)

dat %>% 
  filter(!is.na(rate)) %>%
    ggplot() +
  geom_line(aes(year, rate, group = state),  color = "grey50", 
            show.legend = FALSE, alpha = 0.2, size = 1) +
  geom_line(mapping = aes(year, us_rate),  data = avg, size = 1, color = "black") +
  scale_y_continuous(trans = "sqrt", breaks = c(5, 25, 125, 300)) + 
  ggtitle("Cases per 10,000 by state") + 
  xlab("") + 
  ylab("") +
  geom_text(data = data.frame(x = 1955, y = 50), 
            mapping = aes(x, y, label="US average"), 
            color="black") 
```

3. For the state of California, make a time series plot showing rates for all diseases. Include only years with 10 or more weeks reporting. Use a different color for each disease.

My answer:
```{r}
dat <- us_contagious_diseases %>%
  filter(state == "California" & weeks_reporting > 42) %>%
  mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>% 
  mutate(state = reorder(state, rate)) 

dat %>% 
  filter(!is.na(rate)) %>%
    ggplot() +
  geom_line(aes(year, rate, group = state, col = dat$disease),
            show.legend = FALSE, alpha = 0.7, size = 1) +
  scale_y_continuous(trans = "sqrt", breaks = c(5, 25, 125, 300)) + 
  ggtitle("Cases of California") + 
  xlab("") + 
  ylab("") +
  geom_text(data = data.frame(x = 1955, y = 50), 
            mapping = aes(x, y, label="US average"), 
            color="black") 
```

4. Now do the same for the rates for the US. Hint: compute the US rate by using summarize: the total divided by total population.

My answer:
```{r}
dat <- us_contagious_diseases %>%
  filter(weeks_reporting > 42) %>%
  mutate(rate = sum(count) / sum(population) * 10000 * 52 / weeks_reporting)

dat %>% filter(!is.na(rate)) %>%
  ggplot(aes(year, rate)) +
  geom_line()
```

More thinking needed...








