---
title: "rbasic_20190911_101_kim"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
---

# Chapter 4. Programming basics

By coding in R, we can efficiently perform exploratory data analysis, build data analysis pipelines and prepare data visualization to communicate results.

Advanced R programmers can develop complex packages and even improve R itself.

In this section, we introduce three key programming concepts : conditional expressions, for-loops and functions. These are sometimes useful during data analysis, not just key building blocks for advanced programming.

## 4.1. Conditional expressions

Conditional expressions are used for what is called `flow control`. The most common expression is the if-else statement:
```{r}
a <- 0
if (a!=0){
  print(1/a)
} else{
  print("No reciprocal for 0.")
}
```

The basic idea of this statement is to print the reciprocal of ```a``` unless ```a``` is 0.

```{r}
library(dslabs)
data(murders)
murder_rate <- murders$total / murders$population * 100000
```

The next statement tells us which states, if any, have a murder rate lower than 0.5 per 100,000. **The ```if``` statement protects us from the case in which no state satisfies the condition** :

```{r}
ind <- which.min(murder_rate)

if(murder_rate[ind] < 0.5){
  print(murders$state[ind])
}else{
  print("No state has murder rate that low")
}

if(murder_rate[ind] < 0.25){
  print(murders$state[ind])
}else{
  print("No state has a murder rate that low.")
}
```

A related function that is very useful is ```ifelse```. This function **takes three arguments : a logical and two possible answers**. If the logical is ```TRUE```, the value in the second argument is returned and if ```FALSE```, the value in the third argument is returned:

```{r}
a <- 0
ifelse(a > 0, 1/a, NA)
```

The function is particularly useful because it works on vectors. **It examines each entry of the logical vector and returns elements from the vector provided in the second argument, if the entry is ```TRUE```, or elements from the vector provieded in the third argument, if they entry is ```FALSE```**:

```{r}
a <- c(0, 1, 2, -4, 5)
result <- ifelse(a > 0, 1/a, NA)
result
```

Here is an example of how this function can be radily used to replace all the missing values in a vector with zeros:

```{r}
data(na_example)
no_nas <- ifelse(is.na(na_example), 0, na_example)
#if data in na_example is NA, turns NA into 0.
sum(is.na(no_nas))
#sum of the NAs in no_nas : there's no NA remaining in no_nas
```

Two other useful functions are ```any``` and ```all```. The ```any``` function takes a vector of logicals and returns ```TRUE``` **if any of the entries is ```TRUE```**. The ```all``` function takes a vector of logical and returns ```TRUE``` **if all of the entries are ```TRUE```** :
```{r}
z <- c(TRUE, TRUE, FALSE)
any(z)
all(z)
```

## 4.2. Defining functions

We've computed the average of a vector ```x``` using the ```sum``` and ```length``` functions : ```sum(x)/length(x)```. It is much more efficient to write a function that performs this operation : ```mean(x)```. However, you will encounter situations in which the function does not already exist, so R permits you to write your own :

```{r}
avg <- function(x){
  s <- sum(x)
  n <- length(x)
  s / n
}

x <- 1:100
identical(mean(x), avg(x))
```

Notice that **variables defined inside a function are not saved in the workspace**. So while we use ```s``` and ```n``` when we call ```avg```, the values are created and changed only during the call.

In general, functions are objects, so we assign them to variable names with ```<-```. **The function ```function``` tells R you are about to define a function**.

```{r,error=TRUE}
#general form of function definition
my_function <- function(VARIABLE_NAME){
  perform operations on VARIABLE_NAME and calculate VALUE
  VALUE
}
```

The functions you define can have multiple arguments as well as default values. For example, we can define a function that computes either the arithmetic or geometric average depending on a user defined variable like :

```{r}
avg <- function(x, arithmetic = TRUE){
  n <- length(x)
  ifelse(arithmetic, sum(x)/n, prod(x)^(1/n))
}
```

## 4.3. Namespaces

The `dplyr` function and `stats` package define a ```filter``` function. There are five other examples in `dplyr`. We know this because when we first load `dplyr` we see the following message : 
```{r, error = TRUE}
The following objects are masked from `package:stats`:
  
  filter, lag

The following objects are masked from `package:base`:
  
  intersect, setdiff, setequal, union
```

What does R do when we type ```filter```? Does it use the `dplyr` function or the `stats` function? From our previous work we know it uses the `dplyr` one. But what if we want to use the `stats` version?

These function live in different *namespaces*. R will follow a certain order when searching for a function in these *namespaces*. You can see the order by typing :
```{r}
search()
```
The first entry in this list is the global environment which includes all the objects you define.

If we want to use the `stats` ```filter``` instead of the `dplyr` filter but `dplyr` appears first in the search list, **you can force the use of a specific namespace by using double colons(```::```)** like this :

```{r}
stats::filter
#If wants to be absolutely sure we use the dplyr filter:
dplyr::filter
```

**If we want to use a function in a package without loading the entire package, we can use the double colon as well.**



















