---
title: "tidyverse_20190918_101_kim"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
---

# 5. The tidyverse

We have been manipulating vectors by reordering and subsetting them through indexing. However, once we start more advanced analyses, the preferred unit for data storage is not the vector but **the data frame**, which greatly facilitate the organization of information. We will focus on a specific data format referred to as *tidy* and on specific collection of packages that are particularly helpful for working with *tidy* data referred to as the *tidyverse*.

```{r}
library(tidyverse)
```

We introduce most sidely used tidyverse functionality starting with the **dplyr** package for manipulating data frames and the **purrr** package for working with functions. The tidyverse also includes a graphing package, **ggplot2**, **readr** package, and many others. 

## 5.1. Tidy data

We say that a data table is in *tidy* format **if each row represents one observation and columns represent the different variables available for each of these observations**. The ```murders``` dataset is an example of a tidy data frame.

```{r}
library(dslabs)
head(murders)
```

Each row represent a state with each of the five columns providing a different variable related to these states : name, abbreviation, region, population, and total murders.

For the tidyverse packages to be optimally used, **data need to be reshaped into ```tidy``` format**. 

## 5.2. Exercises










