---
title: "tidyverse_20190923_101_kim"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
---

## 5.7. Summarizing data

An important part of exploratory data analysis is summarizing data. The average and standard deviation are two examples of widely used summary statistics. We cover two new **dplyr** verbs that make computations easier : ```summarize``` and ```group_by```. We learn to access resulting values using the ```pull``` function.

> 5.7.1. ```summarize```

The ```summarize``` function in **dplyr** provides a way to compute summary statistics with intuitive and readable code. Let's start with a simple example based on heights, which includes heights and sex reported by students in an in-class survey:

```{r}
library(dplyr)
library(dslabs)
data(heights)

s <- heights %>% filter(sex == "Female") %>% summarize(average = mean(height), standard_deviation = sd(height))
s
```

This takes our original data table as input, filters it to keep only females, and then produces a new summarized table with just the average and the standard deviation of heights. We get to choose the names of the columns of the resulting table.

As with most other **dplyr** functions, ```summarize``` is aware of the variable names and we can use them directly:

```{r}
heights %>% filter(sex == "Female") %>% summarize(median = median(height), minimum = min(height), maximum = max(height))
```

We can obtain these three values with just one line using the ```quantile``` function : 
```{r, error = TRUE}
heights %>% filter(sex == "Female") %>% summarize(range = quantile(height, c(0, 0.5, 1)))
```

But we will receive an error. **With the function ```summarize```, we can only call functions that return a single value.** 

For another example, let's compute the average murder rate for the United States:

```{r}
data(murders)
murders <- murders %>% mutate(rate = total/population * 100000)

summarize(murders, mean(rate))
```

The US murder rate is not the average of the state murder rates. This is because in the co*mputation above the small states are given the same weight as the large ones. The US murder rate is the total number of murders in the US divided by the total US population :

```{r}
us_murder_rate <- murders %>% summarize(rate = sum(total)/ sum(population) * 10000)

us_murder_rate
```

This computation counts larger states proportionally to their size which results in a larger value.

> 5.7.2. ```pull```

The us_murder_rate object defined above represents just one number. Yet we are storing it in a data frame:

```{r}
class(us_murder_rate)
```

since, as most dplyr functions, summarize always returns a data frame.

This might be problematic if we want to use this result with functions that require a numeric value. Here we show a useful trick for accessing values stored in data when using pipes: when a data object is piped that object and its columns can be accessed using the pull function. To understand what we mean take a look at this line of code:

```{r}
us_murder_rate %>% pull(rate)
```

This returns the value in the rate column of us_murder_rate making it equivalent to us_murder_rate$rate.

To get a number from the original data table with one line of code we can type:

```{r}
us_murder_rate <- murders %>% 
  summarize(rate = sum(total) / sum(population) * 100000) %>%
  pull(rate)

us_murder_rate

class(us_murder_rate)
```

**A common operation in data exploration is to first split data into groups and then compute summaries for each group.** For example, we may want to compute the average and standard deviation for men’s and women’s heights separately. The ```group_by``` function helps us do this.

If we type this:

```{r}
heights %>% group_by(sex)
```

The result does not look very different from ```heights```, except we see ```Groups: sex [2]``` when we print the object. 

Although not immediately obvious from its appearance, this is now a special data frame called a *grouped data frame* and **dplyr** functions, in particular ```summarize```, will behave differently when acting on this object.

Conceptually, you can think of this table as many tables, with the same columns but not necessarily the same number of rows, stacked together in one object. When we ```summarize``` the data after grouping, this is what happens:

```{r}
heights %>% 
  group_by(sex) %>%
  summarize(average = mean(height), standard_deviation = sd(height))
```

The ```summarize``` function applies the summarization to each group separately.

For another example, let’s compute the median murder rate in the four regions of the country:

```{r}
murders %>% 
  group_by(region) %>%
  summarize(median_rate = median(rate))
```

> 5.8.1. Sorting data frames

When examining a dataset, it is often convenient to sort the table by the different columns. We know about the ```order``` and ```sort``` function, but for ordering entire tables, the **dplyr** function arrange is useful. For example, here we order the states by population size:

```{r}
murders %>%
  arrange(population) %>%
  head()
```

With ```arrange``` we get to decide which column to sort by. To see the states by population, from smallest to largest, we arrange by ```rate``` instead:

```{r}
murders %>% 
  arrange(rate) %>% 
  head()
```

Note that the default behavior is to order in ascending order. In **dplyr**, the function ```desc``` transforms a vector so that it is in descending order. To sort the table in descending order, we can type:

```{r}
murders %>% 
  arrange(desc(rate)) %>% 
  head()
```

> 5.8.1. Nested sorting

If we are ordering by a column with ties, we can use a second column to break the tie. Similarly, a third column can be used to break ties between first and second and so on. Here we order by ```region```, then within region we order by murder rate:

```{r}
murders %>% 
  arrange(region, rate) %>% 
  head()
```

> 5.8.2. The top *n*

In the code above, we have used the function ```head``` to avoid having the page fill up with the entire dataset. If we want to see a larger proportion, we can use the ```top_n``` function. This function takes a data frame as it’s first argument, the number of rows to show in the second, and the variable to filter by in the third. Here is an example of how to see the top 10 rows:

```{r}
murders %>% top_n(10, rate)
```

Note that rows are not sorted by ```rate```, only filtered. If want to sort, we need to use ```arrange```. Note that if the third argument is left blank, ```top_n```, filters by the last column.

## 5.9. Exercises

```{r}
library(NHANES)
data(NHANES)
```

1. We will provide some basic facts about blood pressure. First let’s select a group to set the standard. We will use 20-29 year old females. ```AgeDecade``` is a categorical variable with these ages. Note that the category is coded like " 20-29“, with a space in front! What is the average and standard deviation of systolic blood pressure as saved in the ```BPSysAve``` variable? Save it to a variable called ```ref```.

My answer
```{r}
ref <- filter(NHANES, AgeDecade == " 20-29" & Gender == "female") %>% summarize(mean_BP = mean(BPSysAve, na.rm = TRUE), dev_BP = sd(BPSysAve, na.rm = TRUE))

head(ref)
```


2. Using a pipe, assign the average to a numeric variable ```ref_avg```.

My answer
```{r}
ref_avg <- filter(NHANES, AgeDecade == " 20-29" & Gender == "female") %>% summarize(mean_BP = mean(BPSysAve, na.rm = TRUE)) %>% pull(mean_BP)

head(ref_avg)
```

3. Now report the min and max values for the same group.

My answer
```{r}
min(ref)
max(ref)
```

4. Compute the average and standard deviation for females, but for each age group separately rather than a selected decade as in question 1. Note that the age groups are defined by ```AgeDecade```. 

My answer
```{r}
NHANES %>% filter(Gender == 'female') %>% group_by(AgeDecade) %>% summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

5. Repeat exercise 4 for males.

My answer
```{r}
NHANES %>% filter(Gender == "male") %>% group_by(AgeDecade) %>% summarize(mean_ma = mean(BPSysAve, na.rm = TRUE), dev_ma = sd(BPSysAve, na.rm = TRUE))
```

6. We can actually combine both summaries for exercises 4 and 5 into one line of code. This is because ```group_by``` permits us to group by more than one vaqriable. Obtain one big summary table using ```group_by(AgeDecade, Gender)```.

My answer
```{r}
ref_both <- NHANES %>% group_by(AgeDecade, Gender) %>% summarize(mean_ma = mean(BPSysAve, na.rm = TRUE), dev_ma = sd(BPSysAve, na.rm = TRUE))
```

7. For males between the ages of 40-49, compare systolic blood pressure across race as reported in the ```Race1``` variable. Order the resulting table from lowest to highest average systolic blood pressure.

My answer
```{r}
NHANES %>% filter(AgeDecade == " 40-49" & Gender == "male")%>% group_by(Race1) %>% arrange(BPSysAve) %>% head()
```

## 5.10. Tibbles

Tidy data must be stored in data frames. We introduced the ```group_by``` function which permits stratifying data before computing summary statistics. But where is the group information stored in the data frame?

```{r}
murders %>% group_by(region) %>% head()
```

There are no columns with this information. But we can see the line ```A tibble: 6 x 5```. We can learn the class of the returned object using:

```{r}
murders %>% group_by(region) %>% class()
```

The ```tbl```, tibble, is a special kind of data frame. The functions ```group_by``` and ```summarize``` always return this type of data frame. The ```group_by``` function returns a special kind of ```tbl```, the ``grouped_df```. The **dplyr** manipulation verbs(```select```, ```filter```, ```mutate``` and ```arrange```) preserve the class of the input: if they receive a regular data frame they return a regular data frame, while if they receive a tibble they return a tibble. 

But tibbles are the preferred format in the tidyverse and as a result tidyverse functions that produce a data frame from scratch return a tibble. 

Tibbles are very similar to data frames. You can think of them as a modern version of data frames. Nontheless there are three important differences which we describe in the next.

> 5.10.1. Tibbles display better

The print method for tibbles is more readable than that of a data frame. To see this, compare the outputs of typing ```murders``` and the output of murders if we convert it to a tibble. We can do this using ```as_tibble(murders)```. 

> 5.10.2. Subsets of tibbles are tibbles

If you subset the columns of a data frame, you may get back an object that is not a data frame, such as a vector or scalar. **With tibbles this does not happen**:

```{r}
class(as_tibble(murders)[,4])
```

This is useful in the tidyverse since functions require data frames as input.

With tibbles, if we want to access the vector that defines a column, and not get back a data frame, we need to use the accessor ```$```:

```{r}
class(as_tibble(murders)$population)
```

Tibbles will give you a warning if you try to access a column that does not exist:

```{r}
as_tibble(murders)$Population
```

> 5.10.3. Tibbles can have complex entries

While data frame columns need to be vectors of numbers, strings or logical values, tibbles can have more complex objects, such as lists or functions. Also, we can create tibbles with functions:

```{r}
tibble(id = c(1, 2, 3), func = c(mean, median, sd))
```

> 5.10.5. Tibbles can be grouped

The function ```group_by``` returns a special kind of tibble : a grouped tibble. This class stores information that lets you know which rows are in which groups. The tidyverse functions, in particulas the ```summarize``` function, are aware of the group information.

> 5.10.5. Create a tibble using ```data_frame``` instead of ```data.frame```

It is sometimes useful to create our own data frames. To create a data frame in the tibble format, you can do this by using the ```data_frame``` function.

```{r}
grades <- data_frame(names = c("John", "Juan", "Jean", 
                               "Yao"),
                     exam_1 = c(95, 80, 90, 85),
                     exam_2 = c(90, 85, 85, 90))
grades
```

The Base R has a function with a very similar name, ```data.frame```, that can be used to create a regular data frame rather than a tibble. One other important difference is that by default ```data.frame``` coerces characters into factors without providing a warning or message. To avoid this, we use the rather cumbersome argument ```stringAsFactors```:

```{r}
grades <- data.frame(names = c("John", "Juan", "Jean", 
                               "Yao"),
                     exam_1 = c(95, 80, 90, 85),
                     exam_2 = c(90, 85, 85, 90),
                     stringsAsFactors = FALSE)
class(grades$names)
```

To convert a regular data frame to a tibble, we can use the ```as_tibble``` function.

```{r}
as_tibble(grades) %>% class()
```

## 5.11. The dot operator

One of the advantages of using the pipe ```%>%``` is that we do not have to keep naming new objects as we manipulate the data frame:

```{r}
filter(murders, region == "South") %>% 
  mutate(rate = total / population * 10^5) %>% 
  summarize(median = median(rate)) %>%
  pull(median)
```

We can do this because each of these functions takes a data frame as the first argument. But what if we want to access a component of the data frame. For example, what if the ```pull``` function was not available and we wanted to access ```tab_2$rate```? What data frame name would we use? The answer is the dot operator.

```{r}
rates <-filter(murders, region == "South") %>% 
  mutate(rate = total / population * 10^5) %>% 
  .$rate
median(rates)
```

## 5.12. ```do```

The tidyverse functions know how to interpret grouped tibbles. Furthermore, to facilitate stringing commands through the pipe ```%>%```, tidyverse functions consistently return data frames, since this assures that the output of a function is accepted as the input of another. 

But most R functions do not recognize grouped tibbles nor do they return data frames. The ```do``` functions serves as a bridge between R functions such as ```quantile``` and the tidyverse. The ```do``` function understands grouped tibbles and always returns a data frame.

In section 5.7.1, we noted that if we attempt to use ```quantile``` to obtain the min, median and max in one call, we received an error. We can use the ```do``` function to fix this.

First we have to write a function that fits into the tidyverse approach: that is, it receives a data frame and returns a data frame.

```{r}
my_summary <- function(dat){
  x <- quantile(dat$height, c(0, 0.5, 1))
  data_frame(min = x[1], median = x[2], max = x[3])
}
```

But we want a summary for each sex and the code returned just one summary. This is because ```my_summary``` is not part of the tidyverse and does not know how to handled grouped tibbles. ```do``` makes this connection:

```{r}
heights %>% group_by(sex) %>% do(my_summary(.))
```

Here we need to use the dot operator. The tibble created by ```group_by``` is piped to ```do```. Within the call to ```do```, the name of this tibble is ```.``` and we want to send it to ```my_summary```. **If you don't use the dot, then ```my_summary``` has no argument and returns an error telling us that ```argument "dat"``` is missing.**

## 5.13. The purrr package

In section 4.5 we learned about the ```sapply``` function, which permitted us to apply the same function to each element of a vector:

```{r,error=TRUE}
n = 25
compute_s_n <- function(n){
  x <- 1:n
  sum(x)
}
```

and used ```sapply``` to compute the sum of the first ```n``` integers for several values of ```n``` like this:

```{r, error=TRUE}
n <- 1:25
s_n <- sapply(n, compute_s_n)
```

This type of operation, applying the same function or procedure to elements of an object, is quite common in data analysis. The **purrr** package includes functions similar to ```sapply``` but that better interact with other tidyverse functions. The main advantage is that we can better contro l the ouput type of functions. 

In contrast, ```sapply``` can return several different object types : for example, we might expect a numeric result from a line of code, but ```sapply``` might convert our result to character under some circumstances. **purrr** functions will never do this : they will return objects of a specified type or return an error if this is not possible.

The first **purrr** function we will learn is ```map```, which works very similar to ```sapply``` but always, without exception, returns a list:

```{r}
library(purrr)
s_n <- map(n, compute_s_n)
class(s_n)
```

If we want a numeric vector, we can instead use ```map_dbl``` which always returns a vector of numeric values.

```{r}
s_n <- map_dbl(n, compute_s_n)
class(s_n)
```

This produces the same results as the ```sapply``` call shown above.

A particularly useful **purrr** function for interacting with the rest of the tidyverse if ```map_df```, which always returns a tibble data frame. However, the function being called needs to return a vector a or list with names. For this reason, the following code would result in a ```Argument 1 must have names``` error:

```{r}
s_n <- map_df(n, compute_s_n)
```

We need to change the function to make this work:

```{r}
n = 25
compute_s_n <- function(n){
  x <- 1:n
  data_frame(sum = sum(x))
}
s_n <- map_df(n, compute_s_n)
head(s_n)
```

## 5.14. Tidyverse conditionals

> 5.14.1 ```case_when```

The ```case_when``` function is useful for vectorizing conditional statements. It is similar to ```ifelse``` but can output any number of values, as opposed to just ```TRUE``` or ```FALSE```:

```{r}
x <- c(-2, -1, 0, 1, 2)
case_when(x < 0 ~ "Negative", x > 0 ~ "Positive", TRUE ~ "Zero")
```

A common use for this function is to define categorical variables based on existing variables. For example, suppose we want compare the murder rates in three groups of states : New ENgland, West Coast, South, and other. For each state, we need to ask if it is in New England, if it is not we ask if it is in the West Coast, if not we ask if it in the South and if not we assgin other. Here is how we use ```case_when``` to do this:

```{r}
data(murders)
murders %>%
  mutate(group = case_when(
    abb %in%c("ME", "NH", "VT", "MA", "RI", "CT") ~ "New England",
    abb %in% c("WA", "OR", "CA") ~ "West Coast",
    region == "South" ~ "South",
    TRUE ~ "other")) %>%
  group_by(group) %>%
  summarize(rate = sum(total) / sum(population) * 10^5) %>%
  arrange(rate)
```

> 5.14.2. ```between```

A common operation in data analysis is to determine if a value falls inside an interval. We can check this using conditionals. For example to check if the elements of a vector ```x``` are between ``a``` and ```b``` we can type

```{r}
x >= a & x <= b
```

However, this can become cumbersome, especialy within the tidyverse approach. The ```between``` function performs the same operation.

```{r}
between(x, a,b)
```

## 5.15. Exercises

1. Load the ```murders``` dataset. Which of the following is true?

```{r}
data(murders)
View(murders)
```

My answer
B. ```murders``` is in tidy format and is stored in a data frame.

2. Use ```as_tibble``` to covert the ```murders``` data table into a tibble and save it in an object called ```murders_tibble```.

My answer
```{r}
murders_tibble <- as_tibble(murders)
```

3. Use the ```group_by``` function to convert murders into a tibble that is grouped by region.

My answer
```{r}
murders_tibble %>% group_by(region)
```

4. Write tidyverse code that is equivalent to this code:
```exp(mean(log(murders$population)))```. Write it using the pipe so that each function is called without arguments. Use the dot operator to access the population

```{r}
murders %>% .$population %>% log() %>% mean() %>% exp()
```

5. Use the ```map_df``` to create a data frame with three columns named ```n```, ```s_n``` and ```s_n_2```. The first column should contain the numbers 1 through 100. The second and third columns should each contain the sum of 1 through n with n the row number.

```{r}
compute_s_n <- function(n){
  x <- 1:n
  data_frame(sum = sum(x))
}

df_5 <- data.frame("n" = 1:100, "s_n" <- map_df(n, compute_s_n),
                   "s_n_2" <- map_df(n, compute_s_n))
                   
head(df_5)                   
```












