---
title: "The impact of diagnosis age to survival duration"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document:
    latex_engine: xelatex 
---

## 1. Load dataset

```{r message=FALSE}
library(ggplot2, quietly = T)
library(tidyverse, quietly = T)
library(readr, quietly = T)

# wget https://www.dropbox.com/s/xi1ilrp6uzwh4i8/data_20190822_luad_lusc_rnaseq_normalized.RData
load('data_20190822_luad_lusc_rnaseq_normalized.RData')
```

## 2. Processing luad data

```{r}
head(d_luad$days_to_death)
```
```{r}
d1_luad <- d_luad %>% filter(!is.na(days_to_death) & !is.na(age_at_diagnosis))
head(d1_luad$days_to_death)
```

## 2. Visualization

The original purpose was to graphically represent the correlation between the age of cancer and its prognosis. Because I have heard on the Internet that if you get cancer at a young age, the progress of cancer progresses relatively quickly, resulting in more serious prognosis.

Contrary to expectations, this data was not correlated at all. I thought about drawing the plot with another data, but soon I thought it might not be the answer if the shape of the graph appears just as I want. 

So, it's not a planned one, but I've used this graph for visualization.

```{r}
p <- d1_luad %>% ggplot() +
  geom_point(aes(x = age_at_diagnosis/365, y = days_to_death, col = days_to_death)) +
  scale_fill_manual(values=wes_palette(name="GrandBudapest2"))

p
```


